paper

### 2019-01-05 Report/paper.tex
push report changes

    \hypertarget{metadata}{%
\subsection{Metadata}\label{metadata}}

We extract the xml metadata --using the lxmltree python library-- into a
csv file containing the information associated to different tags in its
columns. The provided xml does not exactly follow the xml standard,
hence we handle errors by iteratively parsing the xml tree. We then add
columns that reference the name of the source file and the order in
which an author name occurs in the xml file. From the associated
citation string we add in a shortened version of author names, and the
``author names (year) title'' string that we later use to identify a
paper. Having extracted the metadata we then clean it by unifying names
as slight spelling differences exist within the same author name. This
is done by splitting up names based on whitespace and commas. If the
intersection of the set of substrings of two distinct names contains at
least 2 names of length at least 3 we consider these two names to be
identical. Using this we reduce the number of authors in the dataset
from 1951 down to 1879. To unify the names further, we need to make sure
to not overmerge by labeling to different authors as having the same
name. We do this by constructing the graph of co-authorship and then
only considering names of authors that are in the same neighborhood of a
node on the graph. Then we use the difflib library to check for
similarity of two names and merge them if they reach a the threshold of
0.8. This method again reduces the number of authors in the dataset from
1879 to 1951.

@@

To extract information from the papers themselves we need to first
extract the text from the pdfs. Various parsers performing this task
exist. However, many parsers {[}pdfminer, PyPDF2{]} that claim to be
able to extract the text as well as metadata from papers have not been
kept up to date, which leads to issues with dependencies. We found that
the most reliable parser is poppler https://poppler.freedesktop.org/.

We note that due to inconsistencies with numbering, some of the pdfs
were not extracted correctly from the proceedings. We make sure to
handle these cases.

The advantage of papers is that they are structured fairly regularly.
From ISLS\_Author\_Guidelines\_Oct2014.pdf we know papers the papers in
our dataset should be structured. We use this to our advantage when
extracting information from the paper text. As per
{[}ISLS\_Author\_Guidelines\_Oct2014{]} the document should be
structured in the following way: 1. Title 2. Abstract 3. Keywords 4.
Body of submission (using heading levels 1, 2 and 3 as necessary) 5.
Endnotes (if needed) 6. References 7. Acknowledgements (if needed)

@@

To get the reference section we use a regular expression to find the
`Reference' header and cut of at the end if an Acknowledgement or
Appendix section is present. Python has two modules that implement
regular expressions re and regex. We used regex as it allows to better
handle unicode strings by having unicode expressions, which frequently
occur.

To get the individual reference we tested many ways to split up the
lines, but ultimately, the most robust way was to split up the section
at each new line and the glue the references back together if the lines
satisfy a set of conditions.

The underlying idea is that APA references, the standard referencing
format, all have the same initial structure: Authors (year) title, other
identifying information. We thus check that the beginning of each
reference contains ``Authors (year)''. It is relatively easy to check
for this using regular expressions as Author names and year have a given
structure. If a string does not contain such a section, then it must be
part of an other reference, which is lies before in the list of
reference substrings.

Consider that the references strings are structured the following way
{[}\ldots{}., reference\_part\_i, reference\_part\_i+1, \ldots{}.{]}
Algorithm: 1. Merge names that were cut up. We detect them by checking
against the regular expression: \_\_ 2. Check ending for invalid
sentence endings: reference\_part\_i ends in comma, :, whitespace
uppercase followed by point. If condition holds reference\_part\_i+1 3.
Check beginning for invalid reference starts: not starting in a letter.
Merge with reference\_part\_i if condition holds. 4. Check if a invalid
reference\_part\_i+1 consists only of the authors names. If so Merge
with reference\_part\_i 5. Move up all sentences that do not contain
author names / start of a citation at the beginning 6. Move up all
sentences that do not have a valid citation start

After this we validate the returned split by looking at the overall
length and the initials, as well as checking that a random sample
returns satisfactory splitting accuracy. We find that the above method
of extracting the references produces good results for proper APA
citations and on APA citations outperforms Scolarcy.

    \hypertarget{extracting-institution-country}{%
\subsection{Extracting institution \&
country}\label{extracting-institution-country}}

@@

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping and many
participants use gmail and similar providers as their official
correspondence email we are unable to acquire this information for many
participants. But, Approach 2 may contain differences in spelling in the
institution name, which need to be merged. Thus we combine both
approaches, and use the corresponding institutions in Approach 1 to
unify the institutions obtained in Approach 2.

As institution affiliation does not allow us to automatically associate
a country to the institution, we need to get this through additional
means. A first step is to use the top level domains present in the
emails to associate a country of origin to the participants \& their
associated institutions.

Alongside this, to plot the community we augment the dataset with
geolocation information, having this information we can then easily
associate a country to the institution.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- it was found that crawling
the english, french and german version of Wikipedia (not using the API,
or the API wrapper, as this does not give access to the desired
information) and using beautiful soup the get the coordinates present in
the article was a quick and accurate way to get the location as the
website directs to the relevant article on the institution even when
misspellings are present. But not all institutions, even well known ones
such as Tokyo University, have their coordinates listed on their
Wikipedia. To get the remaining locations we use the geopy, which is a
client for geocoding web services. These results however are more error
prone and will always return a result. Then, to get the country we can
then again use geopy. The resulting data containing paper identifier,
the authors email, it's associated domain, order in which the author was
cited, the institution, country, Longitude and Latitude are saved to
csv. We use the paper identifier and author order to associate this data
with the metadata to link information about authors with the papers they
authored.

@@

\texttt{\color{outcolor}Out[{\color{outcolor}33}]:}

@@

    \hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

@@

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_34_0.png}\end{center}
        \caption{Variation of steady-state oscillation amplitude with driving frequency for an oscillator with undamped frequency of 100 Hz, a driving amplitude $A = 10^6$, and various damping constants.}
        \label{figtest}
    \end{figure}

>>>>

    \hypertarget{processing-the-metadata}{%
\subsection{Processing the Metadata}\label{processing-the-metadata}}

We extract the xml metadata into a csv file containing the information
associated to different tags in its columns using the \emph{lxmltree}
python library. The provided xml does not exactly follow the xml
standard, hence we handle errors by iteratively parsing the xml tree. We
then add columns that reference the name of the source file and the
order in which an author name occurs in the xml file. From the
associated citation string we add in a shortened version of author
names, and the ``author names (year) title'' string that we later use to
identify a paper. Having extracted the metadata we then clean it by
unifying names as slight spelling differences exist within the same
author name. This is done by splitting up names based on whitespace and
commas. If the intersection of the set of substrings of two distinct
names contains at least 2 names of length at least 3 we consider these
two names to be identical. Using this we reduce the number of authors in
the dataset from 1951 down to 1879. To unify the names further, we need
to make sure to not overmerge by labeling to different authors as having
the same name. We do this by constructing the graph of co-authorship and
then only considering names of authors that are in the same neighborhood
of a node on the graph. Then we use the difflib library to check for
similarity of two names and merge them if they reach a the threshold of
0.8. This method again reduces the number of authors in the dataset from
1879 to 1951.

@@

To extract information from the papers themselves we need to first
extract the text from the pdfs. Various parsers performing this task
exist. However, many parsers {[}pdfminer, PyPDF2{]} claiming to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is poppler https://poppler.freedesktop.org/. We
note that due to inconsistencies with numbering, some of the pdfs were
not extracted correctly from the proceedings which leads to text files
that are either empty or do not contain the text of the paper referenced
in the metadata.

The advantage of papers is that they are structured fairly regularly.
From ISLS Guildines \ref{ISLS_Author_Guidelines_Oct2014.pdf} we know
papers the papers in our dataset should be structured. We use this to
our advantage in the following sections when extracting information from
the paper text.

@@

To get the reference section we use a regular expression to find the
`Reference' header and cut of at the end if an Acknowledgement or
Appendix section is present. Python has two modules that implement
regular expressions \emph{re} and \emph{regex}. We used regex as it
allows to better handle unicode strings by having unicode expressions.
This is important as many authors have names that do contain non-ascii
characters. To get the individual reference we tested many ways to split
up the lines, but ultimately, the most robust way was to split up the
section at each new line and the glue the references back together if
the lines satisfy a set of conditions.\\
- The underlying idea is that APA references, the standard referencing
format, all have the same initial structure: Authors (year) title, other
identifying information. We thus check that the beginning of each
reference contains ``Authors (year)''. It is relatively easy to check
for this using regular expressions as Author names and year have a given
structure. If a string does not contain such a section, then it must be
part of an other reference, which is lies before in the list of
reference substrings.

Consider that the references strings are structured the following way
{[}\ldots{}., reference\_part\_i, reference\_part\_i+1, \ldots{}.{]}
Algorithm:

\[
1. Merge names that were cut up. We detect them by checking against the regular expression: __ 
2. Check ending for invalid sentence endings:  reference_part_i ends in comma, :, whitespace uppercase followed by point. If condition holds reference_part_i+1
3. Check beginning for invalid reference starts: not starting in a letter. Merge with reference_part_i if condition holds.
4. Check if a  invalid reference_part_i+1 consists only of the authors names. If so Merge with reference_part_i
5. Move up all sentences that do not contain author names / start of a citation at the beginning 
6. Move up all sentences that do not have a valid citation start
\] After this we validate the returned split by looking at the overall
length and the initials, as well as checking that a random sample
returns satisfactory splitting accuracy. We find that the above method
of extracting the references produces good results for proper APA
citations and on APA citations outperforms Scolarcy.

    \hypertarget{extracting-institution-and-location}{%
\subsection{Extracting Institution and
Location}\label{extracting-institution-and-location}}

@@

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping and many
participants use gmail and similar providers as their official
correspondence email we are unable to acquire this information for many
participants. But, Approach 2 may contain differences in spelling in the
institution name, which need to be merged. Thus we combine both
approaches, and use the corresponding institutions in Approach 1 to
unify the institutions obtained in Approach 2. As institution
affiliation does not allow us to automatically associate a country to
the institution, we need to get this through additional means. A first
step is to use the top level domains present in the emails to associate
a country of origin to the participants \& their associated
institutions. Alongside this, to plot the community we augment the
dataset with geolocation information, having this information we can
then easily associate a country to the institution.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- it was found that crawling
the english, french and german version of Wikipedia (not using the API,
or the API wrapper, as this does not give access to the desired
information) and using \emph{BeautifulSoup} the get the coordinates
present in the article was a quick and accurate way to get the location
as the website directs to the relevant article on the institution even
when misspellings are present. But not all institutions, even well known
ones such as Tokyo University, have their coordinates listed on their
Wikipedia. To get the remaining locations we use the geopy, which is a
client for geocoding web services. These results however are more error
prone and will always return a result. Then, to get the country we can
then again use geopy. The resulting data containing paper identifier,
the authors email, it's associated domain, order in which the author was
cited, the institution, country, Longitude and Latitude are saved to
csv. We use the paper identifier and author order to associate this data
with the metadata to link information about authors with the papers they
authored.

@@

\texttt{\color{outcolor}Out[{\color{outcolor}20}]:}

@@

    \hypertarget{methodology-find-better-section-head..}{%
\section{Methodology {[}find better section
head..{]}}\label{methodology-find-better-section-head..}}

@@

    \hypertarget{results}{%
\section{Results}\label{results}}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_35_0.png}\end{center}
        \caption{Variation of steady-state oscillation amplitude with driving frequency for an oscillator with undamped frequency of 100 Hz, a driving amplitude $A = 10^6$, and various damping constants.}
        \label{figtest}
    \end{figure}
